---
title: "Date exploratory - Group Project"
output: html_document
date: "2025-03-01"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r pressure, echo=FALSE}
library(zoo)
library(lubridate)
library(mgcv)
library(TSA)
library(dynlm)
library(tseries)
library(car)
library(MuMIn)
library(ggplot2)
library(reshape2)
library(greybox)
library(mlr)
library(dplyr)
library(data.table)
library(urca)
library(devtools)
library(prophet)

#Tommy Le: 3/23/2025
library(rugarch)
```

First I load and plot the data, with respective ACF plot.
```{r}
debt_df <- read.csv("Consumer debt payment.csv")
debt <- ts(debt_df$CDSP, start = c(1987, 1), frequency = 4) 
ts.plot(debt, ylab="Consumer debt payment")
acf(debt, lag.max = 4*30)
```

I calulculate spline trend.
```{r}
time.pts = c(1:length(debt))
time.pts = c(time.pts - min(time.pts))/max(time.pts)

spl = gam(debt~s(time.pts)) #Spline
spl_ts = ts(fitted(spl),start=1987,frequency=4)

ts.plot(debt,ylab="homeownership",main="homeownershipwith Trend Estimations")
lines(spl_ts,lwd=2,col="red")
```

I replace outliers at point 134 and 135 with ist trend value.
```{r}
#homeownership[134] <- spl_ts[134]
#homeownership[135] <- spl_ts[135]

#ts.plot(homeownership, ylab="homeownership")
```


```{r}
res <- debt - spl_ts

plot(res, type = "l",main="Residuals Spline detrended")
acf(res, lag.max = 4*30, main='Spline detrended Residuals')
```
There might be some heteroskedasticity.

I clalculate the trend+seasonality.
```{r}
time.pts2 = c(1:length(debt))
time.pts2 = c(time.pts2 - min(time.pts2))/max(time.pts2)

gam_ses = gam(debt~s(time.pts2)+season(debt))
gam_ses_ts = ts(fitted(gam_ses),start=1987,frequency=4)

plot(debt, type = "l",ylab="homeownership",main="homeownership Estimations")
lines(gam_ses_ts, lwd=2, col="red")
```
```{r}
res_2 <- debt - gam_ses_ts

plot(res_2, type = "l",main="Residuals Trend+Seasonality")
acf(res_2, lag.max = 4*50, main='Trend+Seasonality Residuals')
```
Still might be some heterosckedasticity.

I will caluclate the seasonality to the de trended data.
```{r}
model_harmoic=lm(res~harmonic(res))
summary(model_harmoic)
```


I differenciate the data.
```{r}
diff_ts = diff(ts(debt,start=1987,frequency=4))

plot(diff_ts, type = "l",main="Residuals Differencing")
acf(diff_ts, lag.max = 4*30, main='Differencing Residuals')
```
I differenciate again.

```{r}
diff_diff_ts = diff(ts(diff_ts,start=1987,frequency=4))

plot(diff_diff_ts, type = "l",main="Residuals Diff_Diff")
acf(diff_diff_ts, lag.max = 4*30, main='Diff_Diff Residuals')
```

Start using ARIMA.
```{r}
n <- length(res_2)
p <- 0:6
q <- 0:6
aic <- matrix(0, length(p), length(q)) 
for (i in 1:length(p)) {
  for (j in 1:length(q)) {
    modij = arima(res_2, order = c(p[i], 0, q[j]), method = 'ML')
    aic[i, j] = modij$aic + 2 * (p[i] + q[j] + 1) * (p[i] + q[i]) / (n - p[i] - q[j] - 1)
  }
}


norder = 7
aicv = as.vector(aic)
plot(aicv, ylab = "AIC values")
indexp = rep(c(1:norder),norder)
indexq = rep(c(1:norder),each=norder)
indexaic = which(aicv == min(aicv))
porder = indexp[indexaic]-1
qorder = indexq[indexaic]-1
cat("The p order is ", porder)
cat("\nThe q order is ", qorder)
```
```{r}
res_model = arima(res_2, order = c(3, 0, 6), method = 'ML')
res_model
```
```{r}
ts.plot(residuals(res_model),ylab = "Residual Process")
acf(residuals(res_model),lag.max = 4*30)
pacf(residuals(res_model),lag.max = 4*30)
qqPlot(residuals(res_model), ylab = "Residuals")
Box.test(res_model$resid, lag = (3 + 6 + 1), type = "Ljung-Box", fitdf = (3+6))
```






```{r}

test_modelA <- function(p,d,q){
  mod = arima(debt, order=c(p,d,q), method="ML", include.mean = TRUE)
  current.aic = AIC(mod)
  df = data.frame(p,d,q,current.aic)
  names(df) <- c("p","d","q","AIC")
  #print(paste(p,d,q,current.aic,sep=" "))
  return(df)
}

orders = data.frame(Inf,Inf,Inf,Inf)
names(orders) <- c("p","d","q","AIC")


for (p in 0:8){
  for (d in 0:1){
    for (q in 0:8) {
      possibleError <- tryCatch(
        orders<-rbind(orders,test_modelA(p,d,q)),
        error=function(e) e
      )
      if(inherits(possibleError, "error")) next
      
    }
  }
}
orders <- orders[order(orders$AIC),]
orders

```
```{r}
model <- arima(debt, order = c(2,1,5), method = "ML", include.mean = TRUE)
model
```

```{r}
ts.plot(residuals(model),ylab = "Residual Process")
acf(residuals(model),lag.max = 4*30)
pacf(residuals(model),lag.max = 4*30)
qqPlot(residuals(model), ylab = "Residuals")
Box.test(model$resid, lag = (2 + 1 + 5 + 1), type = "Ljung-Box", fitdf = (2 + 1 + 5))
```
```{r}
#add garch
# Fit the ARIMA model
model <- arima(debt, order = c(2, 1, 5), method = "ML", include.mean = TRUE)

# Check residuals of the ARIMA model
ts.plot(residuals(model), ylab = "Residual Process")
acf(residuals(model), lag.max = 4*30)
pacf(residuals(model), lag.max = 4*30)
qqPlot(residuals(model), ylab = "Residuals")
Box.test(model$resid, lag = (2 + 1 + 5 + 1), type = "Ljung-Box", fitdf = (2 + 1 + 5))

# Fit a GARCH model on residuals of the ARIMA model
garch_model <- ugarchfit(spec = ugarchspec(variance.model = list(model = "sGARCH", garchOrder = c(1,1)),
                                           mean.model = list(armaOrder = c(2, 5), include.mean = TRUE)),
                         data = residuals(model))

# Display the summary of the GARCH model
summary(garch_model)

# Calculate MAPE for ARIMA model
predictions <- fitted(model)  # Predictions from the ARIMA model
actual <- debt  # Actual data

mape <- mean(abs((actual - predictions) / actual)) * 100  # MAPE formula
cat("The MAPE is:", mape, "%")
```
