---
title: "Date exploratory - Group Project"
output: html_document
date: "2025-03-01"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r pressure, echo=FALSE}
library(zoo)
library(lubridate)
library(mgcv)
library(TSA)
library(dynlm)
library(tseries)
library(car)
library(MuMIn)
library(ggplot2)
library(reshape2)
library(greybox)
library(mlr)
library(dplyr)
library(data.table)
library(urca)
library(devtools)
library(prophet)
library(rugarch)
```

First I load and plot the data, with respective ACF plot.
```{r}
mortage_df <- read.csv("Mortgage Debt Service Payments as a Percent of Disposable Personal Income.csv")
mortage <- ts(mortage_df$MDSP, start = c(1987, 1), frequency = 4) 
ts.plot(mortage, ylab="Consumer debt payment")
acf(mortage, lag.max = 4*30)
```
Very similar plot and ACF plot to consumer debt payment...


I calulculate spline trend.
```{r}
time.pts = c(1:length(mortage))
time.pts = c(time.pts - min(time.pts))/max(time.pts)

spl = gam(mortage~s(time.pts)) #Spline
spl_ts = ts(fitted(spl),start=1987,frequency=4)

ts.plot(mortage,ylab="Mortage Debt Payment",main="Mortage Debt Payment Trend Estimations")
lines(spl_ts,lwd=2,col="red")
```


```{r}
res <- mortage - spl_ts

plot(res, type = "l",main="Residuals Spline detrended")
acf(res, lag.max = 4*30, main='Spline detrended Residuals')
```

I clalculate the trend+seasonality.
```{r}
time.pts2 = c(1:length(mortage))
time.pts2 = c(time.pts2 - min(time.pts2))/max(time.pts2)

gam_ses = gam(mortage~s(time.pts2)+season(mortage))
gam_ses_ts = ts(fitted(gam_ses),start=1987,frequency=4)

plot(mortage, type = "l",ylab="Mortage Debt Payment",main="Mortage Debt Payment Estimations")
lines(gam_ses_ts, lwd=2, col="red")
```
```{r}
res_2 <- mortage - gam_ses_ts

plot(res_2, type = "l",main="Residuals Trend+Seasonality")
acf(res_2, lag.max = 4*50, main='Trend+Seasonality Residuals')
```
No improvement.

I differenciate the data.
```{r}
diff_ts = diff(ts(mortage,start=1987,frequency=4))

plot(diff_ts, type = "l",main="Residuals Differencing")
acf(diff_ts, lag.max = 4*30, main='Differencing Residuals')
```
I differenciate again.

```{r}
diff_diff_ts = diff(ts(diff_ts,start=1987,frequency=4))

plot(diff_diff_ts, type = "l",main="Residuals Diff_Diff")
acf(diff_diff_ts, lag.max = 4*30, main='Diff_Diff Residuals')
```

Start using ARIMA on just the detrended data as it is simpler.
```{r}
n <- length(res)
p <- 0:6
q <- 0:6
aic <- matrix(0, length(p), length(q)) 
for (i in 1:length(p)) {
  for (j in 1:length(q)) {
    modij = arima(res, order = c(p[i], 0, q[j]), method = 'ML')
    aic[i, j] = modij$aic + 2 * (p[i] + q[j] + 1) * (p[i] + q[i]) / (n - p[i] - q[j] - 1)
  }
}


norder = 7
aicv = as.vector(aic)
plot(aicv, ylab = "AIC values")
indexp = rep(c(1:norder),norder)
indexq = rep(c(1:norder),each=norder)
indexaic = which(aicv == min(aicv))
porder = indexp[indexaic]-1
qorder = indexq[indexaic]-1
cat("The p order is ", porder)
cat("\nThe q order is ", qorder)
```
p equal to 3 and q to 6 are the same values of costumer debt payment.
```{r}
res_model = arima(res, order = c(3, 0, 6), method = 'ML')
res_model
```
```{r}
ts.plot(residuals(res_model),ylab = "Residual Process")
acf(residuals(res_model),lag.max = 4*30)
pacf(residuals(res_model),lag.max = 4*30)
qqPlot(residuals(res_model), ylab = "Residuals")
Box.test(res_model$resid, lag = (3 + 6 + 1), type = "Ljung-Box", fitdf = (3+6))
```






```{r}

test_modelA <- function(p,d,q){
  mod = arima(mortage, order=c(p,d,q), method="ML", include.mean = TRUE)
  current.aic = AIC(mod)
  df = data.frame(p,d,q,current.aic)
  names(df) <- c("p","d","q","AIC")
  #print(paste(p,d,q,current.aic,sep=" "))
  return(df)
}

orders = data.frame(Inf,Inf,Inf,Inf)
names(orders) <- c("p","d","q","AIC")


for (p in 0:8){
  for (d in 0:1){
    for (q in 0:8) {
      possibleError <- tryCatch(
        orders<-rbind(orders,test_modelA(p,d,q)),
        error=function(e) e
      )
      if(inherits(possibleError, "error")) next
      
    }
  }
}
orders <- orders[order(orders$AIC),]
orders

```
```{r}
model <- arima(mortage, order = c(1,1,2), method = "ML", include.mean = TRUE)
model
```

```{r}
ts.plot(residuals(model),ylab = "Residual Process")
acf(residuals(model),lag.max = 4*30)
pacf(residuals(model),lag.max = 4*30)
qqPlot(residuals(model), ylab = "Residuals")
Box.test(model$resid, lag = (1 + 1 + 2 + 1), type = "Ljung-Box", fitdf = (1 + 1 + 2))
```
These are much better results.

```{r}

arima_residuals <- residuals(model)

garch_spec <- ugarchspec(variance.model = list(model = "sGARCH", garchOrder = c(1, 1)),
                         mean.model = list(armaOrder = c(1, 2)),
                         distribution.model = "std")
garch_model <- ugarchfit(spec = garch_spec, data = arima_residuals)
garch_model
garch_forecast <- ugarchforecast(garch_model, n.ahead = 4)
garch_fitted_values <- fitted(garch_forecast)
garch_fitted_values
# Calculate the MAPE (Mean Absolute Percentage Error)
actual_values <- window(mortage, start = c(2023, 3), end = c(2024, 4))
length(actual_values)
length(garch_fitted_values)
mape <- mean(abs((actual_values - garch_fitted_values) / actual_values)) * 100
cat("MAPE: ", mape, "%\n")


```
